Contextual Identity: Towards Usable Privacy

### 2013-03-01 paper.tex
more reorg

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. It is
also true that in offline environments, humans don't have perfect memories and
will often eventually forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online environments~\cite{delete}. Managing contextual identities
online has become increasingly difficult and fraught with mistakes; we must
provide users with better tools to control their contextual identities.

@@

\section{Contextual identity violations}
We consider it a contextual identity violation when
multiple contextually-separate identities are linked without the user's
permission, or when the user cannot select which contextual identity is
presented to a given audience.

@@

\subsection{How do users think about identity?}
Users' contextual identities may be task-oriented, service-oriented,
relationship-oriented, or guided by other principles. The long-time existence
of support for multiple accounts in email clients suggests that many users have
multiple email identifiers, which could be considered as a proxy for identity
if tied to service accounts. Similarly, data suggests that a large minority of
Twitter users have multiple Twitter accounts, and the plethora of cross-posting
or syndication software from one social network to another suggests that many
users take advantage of multiple social networks~\cite{twitter,xpost}.  That is
not to say that all users have the same tolerance or even need for multiple
explicit identities: in fact many users may be exclusively task-oriented and
use the Internet with specific tasks in mind rather than for social interaction
or exploratory surfing\footnote{Even these users are subject to building
implicit identities through the use of tracking techniques, however.}.

\subsection{How do users manage identity?}
\paragraph{Profile management}
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult
to use that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.} Tang et
al.~show that most people do not take the time to create labels and maintain
friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

\paragraph{Private browsing mode}
Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, the threat
model for using private browsing mode may be other household members, or it may
be a service to which the user wants to limit cookie leakage.

@@

\paragraph{Mitigating information leaks}
Given a user's contextual identities, we may be able to prevent information
leakage between them.  For example, preventing third-party cookies from being
read and set across task boundaries means that targeted ads for dating would
not follow the user when visiting a technology news site. Knowing the user's
current contextual identity also allows the browser to suppress URL bar
suggestions that aren't relevant to that identity, protecting the user from
both shoulder-surfing and distraction.

\paragraph{Mitigating unwanted linkage}
Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\subsection{Expiring contextual identities}
Oftentimes a user may have a long-term task, such as wedding planning or house
buying, that requires constructing a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
such as cookie data.

\subsection{Recovering from Mistakes}
\subsubsection{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, a frequent cause
of regret~\cite{wang}.  Although this approach does not help when the service
provider decides to change the audience on the user's behalf, auditing gives
the user a good opportunity to review posts that might cause regrets. In the
case of high-volume users, such a tool could incorporate sentiment analysis to
aid the user in finding particularly problematic content.

@@

\subsubsection{Remembering to Forget}
Humans don't have perfect memories, and perhaps social networks should not
either~\cite{delete}. Many users already engage in manual auditing and deletion
of old posts~\cite{fbtips2}. A better solution would be to build tools that let
the user manage this more easily through an API: both Twitter and Facebook
provide APIs for deletion\footnote{At the time of this writing, Google Plus
offers a read-only API.}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

\subsection{Preventing Mistakes}
\subsubsection{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience.

\subsubsection{Visual reminders of context}
To mitigate privacy violations caused by poor user interfaces, the onus must be
on software developers to provide sufficient visual cues of context. For
example, many but not all Google products support multiple accounts. However,
the visual cue to indicate which account is active is a small, easy to ignore
text name at the top right of the page.

>>>>

In offline environments, managing contextual identities is more intuitive than
in online environments. There are cues to help you determine where you are,
who your intended audience is, how many people will overhear you, and how
likely your information is to be re-broadcast in a different context. 
In offline environments, humans don't have perfect memories and
will typically forget.  However, with the advent of increasingly vast
searching, indexing, and archiving capabilities, one cannot rely on forgetting
in online environments~\cite{delete}. Managing contextual identities
online has become increasingly difficult and fraught with mistakes; we must
provide users with better tools to control their contextual identities.

@@

\section{Contextual identity violations}
We consider it a contextual identity violation when multiple identities are
linked without user intent, or when the user cannot choose which contextual
identity to assert.

@@

\subsection{How do users think about identity?}
It is unlikely that the majority of users are consciously aware of having
multiple contextual identities. Imagine two users at the opposite ends of the
spectrum:
\begin{itemize}
\item A user who participates in no social networks, reads paper newspapers,
uses the Internet for well-defined tasks (e.g., coordinating visits with
friends over email, looking up restaurant reviews) but does not otherwise
engage in discretionary browsing.
\item A savvy user who is active on Facebook, Twitter, Instagram and has
different posting levels, topics, and followers on each, maintains multiple
email accounts and pseudonyms, is online multiple hours daily for both work and
personal reasons. 
\end{itemize}

The second user may be much more aware of contextual identities, but even the
first user would benefit from protection against contextual identity
violations. For example, engaging in a long-term task such as house-buying
often requires the construction of a contextual identity. The tasks associated
with this identity have an externally imposed end date and may have long-lived
side-effects, such as long-term cookies that reveal sensitive information
(e.g., the neighborhood of the house purchase, income information derived from
purchases, other demographic data). For contextual identities that have
outlived their usefulness, users would benefit from destruction of side-effects
such as cookie data.

Users' contextual identities may be task-oriented, service-oriented,
relationship-oriented, or guided by other principles. The long-time existence
of support for multiple accounts in email clients suggests that many users have
multiple email identifiers, which could be considered as a proxy for identity
if tied to service accounts. Similarly, data suggests that a large minority of
Twitter users have multiple Twitter accounts, and the plethora of cross-posting
or syndication software from one social network to another suggests that many
users take advantage of multiple social networks~\cite{twitter,xpost}.  That is
not to say that all users have the same tolerance or even need for multiple
explicit identities: in fact many users may be exclusively task-oriented and
use the Internet with specific tasks in mind rather than for social interaction
or exploratory surfing\footnote{Even these users are subject to building
implicit identities through the use of tracking techniques, however.}.

\subsection{How do users manage identity?}
\paragraph{Private browsing mode}
Privacy modes in Firefox, Chrome, Safari, and Opera are a good first step, but
are no means a complete solution~\cite{ABBJ10}. There is no standard behavior
for private browsing mode, and thus interaction with extensions, treatment of
cookies, history, and bookmarks upon entering and exiting this mode are
different and serve different use cases. Depending on user desire, the threat
model for using private browsing mode may be other household members, or it may
be a service to which the user wants to limit cookie leakage.

\paragraph{Profile management}
Although a user may subconsciously be aware of their contextual identities,
expecting users to enumerate their identities and manually curate them is
probably too burdensome. Firefox has a Profile Manager, but it is so difficult
to use that it will be removed.\footnote{See
\url{https://bugzilla.mozilla.org/show\_bug.cgi?id=214675\#c53}.} Tang et
al.~show that most people do not take the time to create labels and maintain
friend lists~\cite{tang}. However, the same authors show how to use
heterogeneous networks to infer relationship types (e.g., manager-subordinate)
though the graph itself is unlabeled.

@@

\paragraph{Mitigating implicit linkage}
Given a user's contextual identities, we may be able to prevent information
leakage between them.  For example, preventing third-party cookies from being
read and set across task boundaries means that targeted ads for dating would
not follow the user when visiting a technology news site. Knowing the user's
current contextual identity also allows the browser to suppress URL bar
suggestions that aren't relevant to that identity, protecting the user from
both shoulder-surfing and distraction.

\paragraph{Mitigating explicit unwanted linkage}
Another partial solution is to prompt the user when they're about to link
multiple contextual identities through a federated login service. For example,
Persona is the suite of tools that implements the BrowserID protocol, and it
supports multiple email identifiers~\cite{browserid}.
Because we can categorize the type of service by URL
as \texttt{about:profile} does, Persona could prompt the user when they're about
to link the same email identifier to radically different contextual identities,
such as ones for dating and professional use.

\paragraph{Controlling Information Flow}
Sometimes a user might want to prevent mistakes from being made in the first
place. If a user is concerned about a particular service provider, they can
install an extension like Disconnect to disallow cookies from that provider
unless they explicitly want to interact with that site~\cite{disconnect}.
This type of tool (cookie blocker) helps users assert control over audience.

\paragraph{Auditing}
A major cause of user embarrassment is exposing information to the wrong
audience, as illustrated in section~\ref{sec:examples}. A browser is in an
ideal position to intermediate social network posts, aggregate them locally,
and present them to the user when they desire to audit their digital footprint.
For example, a user could see all of the comments and posts they made in the
last week and redact content that is sensitive or too negative, a frequent cause
of regret~\cite{wang}.  Although this approach does not help when the service
provider decides to change the audience on the user's behalf, auditing gives
the user a good opportunity to review posts that might cause regrets. In the
case of high-volume users, such a tool could incorporate sentiment analysis to
aid the user in finding particularly problematic content.

@@

\paragraph{Remembering to Forget}
Humans don't have perfect memories, and perhaps social networks should not
either~\cite{delete}. Many users already engage in manual auditing and deletion
of old posts~\cite{fbtips2}. A better solution would be to build tools that let
the user manage this more easily through an API: both Twitter and Facebook
provide APIs for post deletion\footnote{Google Plus
offers a read-only API.}. One way to reduce the likelihood of linked contextual
identities would be to provide applications that expire posts after some time.

\paragraph{Visual reminders of context}
To mitigate privacy violations caused by poor user interfaces, the onus must be
on software developers to provide sufficient visual cues of context. For
example, many but not all Google products support multiple accounts. However,
the visual cue to indicate which account is active is a small, easy to ignore
text name at the top right of the page.
